### 2016-06-25 testbed.tex
worked on testbed

# Attempt to fetch new configuration files
if ! wget -O /tmp/${config_file} ${url%/}/${config_file}; then
    echo "Warning: Unable to update rasp_config files"
    exit 1
fi

# Unzip and overwrite configurationn files to root's home directory
unzip -q -o ${config_file} -d /home/pi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsection{Installation and Updating the image}

@@

% CHROOT to OS image
\begin{lstlisting}[]
$ cd $ROOTDIR
# cp /etc/resolv.conf etc/resolv.conf
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

% CHROOT to OS image
\begin{lstlisting}[]
# cp /usr/bin/qemu-arm-static usr/bin
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

The bootloader should be made aware of the initramfs that we will generate:
%Raspbian lite uses "u-boot" to boot, so it should be made aware of the
%initramfs that we will generate.

% Tell uboot to load initramfs
\begin{lstlisting}[]
(chroot) # echo "initramfs init.gz" > /boot/config.txt
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, lets generate the image. This is done using "mkinitramfs", that will
by default look for the kernel modules of the current system. Remember we are
chrooted, so we need to specify the correct kernel modules to look for. These
can be found in "/lib/modules".

% Create an initramfs
\begin{lstlisting}[]
(chroot) # ls /lib/modules/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

For \ac{Raspi} 1, our kernel modules was "4.4.11+". The other modules in /lib/modules
is for \ac{Raspi} 2.

Thus, to run mkinitramfs, type in the correct kernel version:

% Create an initramfs
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.11+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Although there might be printet a few warnings, this should generate
a new working \texttt{initramfs} that will put a \ac{RAM} filesystem
on top of the root filesystem.

At this point, it is no longer required to be \texttt{chrooted} and the image
can be written to a memory card. The following commands exits the \texttt{chrooted}
environment, unmount all partitions, and detaches the loop device:

@@

\subsection{(Optional) Persistent subdirectories}

@@

>>>>

# Attempt to fetch new configuration files
if ! wget -O -q --show-progress /tmp/${config_file} ${url%/}/${config_file}; then
    echo "Warning: Unable to update rasp_config files"
    exit 1
fi

# Unzip and overwrite configurationn files to root's home directory
unzip -q -o /tmp/${config_file} -d /home/pi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsection{Installing and updating the image}

@@

% CHROOT to OS image
\begin{lstlisting}[]
$ cd $ROOTDIR
# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

% CHROOT to OS image
\begin{lstlisting}[]
# cp /usr/bin/qemu-arm-static ${ROOTDIR}/usr/bin
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

%Finally, lets generate the initial \ac{RAM} filesystem. This is done using the command
%\texttt{mkinitramfs}, that will by default look for the kernel modules of the current system. Remember we are
%chrooted, so we need to specify the correct kernel modules to look for. These
%can be found in "/lib/modules".
Now, the initial \ac{RAM} filesystem can be generated. \ac{Raspi} version 1
needs a different kernel than \ac{Raspi} version 2 and version 3. Thus, start
by finding the available kernel modules:
%Because \texttt{initramfs} is executed within a
%\texttt{chrooted} environment, it will not be able to detect the
%kernel modules in the \texttt{chrooted} environment. The kernel modules will

% Create an initramfs
\begin{lstlisting}[]
(chroot) # ls /lib/modules/
4.4.13+  4.4.13-v7+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Kernel "4.4.13+" is for \ac{Raspi} version 1 and kernel "4.4.13-v7+" works
for \ac{Raspi} version 2 and version 3.

Lets proceed by generating an initial \ac{RAM} filesystem for both kernels:

%Lets proceed with kernel modules "4.4.13+".
%Lets create an initial ram filesystem for both kernels.

%Thus, to run mkinitramfs, type in the correct kernel version:

% Create an initramfs
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+

Although there might be printet a few warnings, the above commands should successfully
generate working initial \ac{RAM} filesystems.

An initial \ac{RAM} filesystem needs to be called by the bootloader. In Raspbian
this is done by adding a command to \texttt{config.txt} in the boot partition.

If the system should be run in a \ac{Raspi} version 1, then use \texttt{init.gz}:
%The bootloader should be made aware of the initramfs that we will generate:
\begin{lstlisting}[]
(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # Only for Raspberry Pi version 1
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the system should be run in a \ac{Raspi} version 2 or 3, then use \texttt{init-v7.gz}:
\begin{lstlisting}[]
(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # Only for Raspberry Pi version 2 and 3
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

At this point, it is no longer required to be \texttt{chrooted} and the image
can be written to a memory card. The following commands exits the \texttt{chrooted}
environment, unmount all partitions, and detaches the loopback device:

@@

\subsection{(Optional) Persistent subdirectories for distributed Linux image}

EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
WRITTEN TO THE MEMORY CARDS

@@

EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
WRITTEN TO THE MEMORY CARDS

### 2016-06-25 testbed_http.tex
worked on testbed

\begin{lstlisting}[]
$ export HTTPBOOTIMAGE="${WORKDIR}/httpboot"
$ export HTTPROOTIMAGE="${WORKDIR}/httproot"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Extract root partition from image
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
# dd if=${DEV}p1 of=${HTTPBOOTIMAGE}.img bs=4M
# dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Mount http root and home directories (ADD LOOP OPTION HERE WHEN MOUNTING)
\begin{lstlisting}[]
$ export HTTPROOTDIR="${WORKDIR}/httproot"
$ mkdir -p ${HTTPROOTDIR}
# mount ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
# mount ${HTTPBOOTIMAGE}.img ${HTTPROOTDIR}/boot
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

#. /scripts/functions # Already loaded in the init script

@@

    # If the image is not cached, then get it
    if [ ! -f ${img} ] && wget --spider -q "${URL}/${img}"; then
        # Delete old image
        rm -f $(readlink -f ${IMAGEPATH})

@@

Add the below to \$\{HTTPROOTDIR\}/boot/cmdline.txt (CHANGE URL HERE) (remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)
\begin{lstlisting}[]
boot=http ro ip=dhcp image=/dev/mmcblk0p2:/root.img root=/dev/loop0 url=http://192.168.87.106/raspi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% ip=dhcp root=/dev/http http-options=hard,intr,ro

Edit fstab to look like this (exclude the last line if persistent
home storage directory is not desired):
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/fstab}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
proc            /proc           proc    defaults          0       0
/dev/mmcblk0p1  /boot           vfat    defaults          0       2
/dev/loop0      /               ext4    defaults,noatime  0       1
/dev/mmcblk0p2  /home           ext4    defaults,noatime  0       0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

Create new initramfs (remember that your kernel might be different)
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.11+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

Unmount:
\begin{lstlisting}[]
# umount ${HTTPROOTIMAGE}/{dev/pts,proc,sys,dev}  
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

Mount the boot partition from the memory card. Copy the content from the
boot image and unmount it again:
\begin{lstlisting}[]
# mount ${HTTPDEV}p1 ${HTTPROOTDIR}/mnt
# cp -r ${HTTPROOTDIR}/boot/* ${HTTPROOTDIR}/mnt/ && sync
# umount ${HTTPROOTDIR}/mnt
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Mount the home partition from the memory card. Copy the content from the
root image and unmount it again:
\begin{lstlisting}[]
# mount ${HTTPDEV}p2 ${HTTPROOTDIR}/mnt
# cp -r ${HTTPROOTDIR}/home/* ${HTTPROOTDIR}/mnt/ && sync
# umount ${HTTPROOTDIR}/mnt
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

>>>>

%This part will explain how to prepare the existing image to be downloaded from
%a \ac{HTTP} server.
The Linux image that has been made so far has two partitions. One 
partition with the boot files and one containing the root file system.

This section will present how the partitions can be extracted from the image,
and how the boot partition can be configured to download the root filesystem
from a \ac{HTTP} server during startup.

The advantage of this setup is that multiple \ac{Raspi}s can share the same
root image. This significantly eases the maintainance task.

%An alternative to the setup explained in this procedure is to use \ac{NFS}
%instead of \ac{HTTP}.

%The first program called when a \ac{Raspi} starts is the Boot Loader,
%\texttt{U-Boot}. \texttt{U-Boot} 
%
%When Raspbian starts up, the first program called
%
%The procedure explained below will customize the initial \ac{RAM} filesystem
%
%
%there exists only one root file
%
%The advantage of this is that the root filesystem only one all \ac{Raspi}s can automatically download the 
%
%The boot partition will be required in the \ac{Raspi}s, but the root image can

Start by declaring the path and names of the boot- and root-image:. 
\begin{lstlisting}[]
$ export HTTPBOOTIMAGE="${WORKDIR}/httpboot"
$ export HTTPROOTIMAGE="${WORKDIR}/httproot"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The, setup a loopback device of the existing image and use \texttt{dd} to
extract the boot and root partitions from the image:
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
# dd if=${DEV}p1 of=${HTTPBOOTIMAGE}.img bs=4M
# dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
# losetup -d ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now it is time to configure the boot- and root-image just extracted before
the content of the boot image is copied to the a memory card and the root
image is uploaded to a \ac{HTTP} server.

Create a mountpoint and mount the http root and boot images:
%Mount the http root image and home directories
\begin{lstlisting}[]
$ export HTTPROOTDIR="${WORKDIR}/httproot"
$ mkdir -p ${HTTPROOTDIR}
# mount -o loop ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
# mount -o loop ${HTTPBOOTIMAGE}.img ${HTTPROOTDIR}/boot
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

In a previous steps, the boot loader was configured to call an initial \ac{RAM}
filesystem. The coming steps will present how to the initial \ac{RAM}
filesystem can be extended further to download and cache the root filesystem
from a \ac{HTTP} server before mounting it.

The initial \ac{RAM} filesystem contains an \texttt{init} script that
sets up the prepares the yis first
called. This script initializes the  may be rewritten, but

@@

@@

    # If the image is not cached, then get it
    if [ ! -f "${IMAGEPATH%/*}/${img}" ] && wget --spider -q "${URL}/${img}"; then
        # Delete old image
        rm -f $(readlink -f ${IMAGEPATH})

@@

Keep the content in \texttt{cmdline.txt}, but add/modify the variables
\texttt{boot}, \texttt{ip}, \texttt{image}, \texttt{root}, and \texttt{url}.

%Keep also all other variables as they are in
%\texttt{cmdline.txt}.

%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
(remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
<@\Suppressnumber @>boot=http ip=dhcp image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0
    url=http://kom.aau.dk/project/TuneSCode/raspi/<@\Reactivatenumber @>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% ip=dhcp root=/dev/http http-options=hard,intr,ro

Here is how the variables are used in the initramfs:
\begin{itemize}
    \item boot=http: tells the init script in initramfs to call the \texttt{http} script presented above
    \item ip=dhcp: use \ac{DHCP} to obtain an \ac{IP} address in the initial ram filesystem
    \item image=<device>:<root image path>: is where the \ac{Raspi} will store the downloaded root image.
    \item root=/dev/loop0: The \texttt{http} initramfs script will make the root filesystem available on /dev/loop0
    \item url=<URL to root image>: Is the directory on a \ac{HTTP} server where the root image will be iavailable.
\end{itemize}

%The \ac{URL} should be modified to reflect where
%the \ac{Raspi}s should download the http root image. The \texttt{image} variable informs the
%\texttt{http} initramfs script to download and cache the http root image in the second partition
%on its memory card at path \texttt{/httproot.img}.

Edit fstab to look like this (exclude the last line if a persistent
home storage directory is not desired):
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/fstab}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
proc            /proc           proc    defaults          0       0
/dev/mmcblk0p1  /boot           vfat    defaults          0       2
/dev/loop0      /               ext4    defaults,noatime  0       1
/dev/mmcblk0p2  /home           ext4    defaults,noatime  0       0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

Create new initramfs (remember that your kernel might be different)
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

Unmount:
\begin{lstlisting}[]
$ cd ..
# umount ${HTTPROOTIMAGE}/{dev/pts,proc,sys,dev}  
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

Mount the boot partition from the memory card. Copy the content from the
boot image and unmount it again:
\begin{lstlisting}[]
# mount ${HTTPDEV}p1 ${HTTPROOTDIR}/mnt
# cp -rp ${HTTPROOTDIR}/boot/* ${HTTPROOTDIR}/mnt/ && sync
# umount ${HTTPROOTDIR}/mnt
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Mount the home partition from the memory card. Copy the content from the
root image and unmount it again:
\begin{lstlisting}[]
# mount ${HTTPDEV}p2 ${HTTPROOTDIR}/mnt
# cp -rp ${HTTPROOTDIR}/home/* ${HTTPROOTDIR}/mnt/ && sync
# umount ${HTTPROOTDIR}/mnt
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\begin{itemize}
    \item Copy http root image (\$\{HTTPROOTIMAGE\}.img) a HTTP server
    \item Create images.csv
\end{itemize}

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{http://<SERVERIP>/.../images.csv}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
# Ethernet MAC    Root image
b8:27:eb:5b:da:20 httproot_1.img
b8:27:eb:7b:c3:91 httproot_1.img
b8:27:eb:7b:c3:91 httproot_2.img
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Notice that the \ac{Raspi}s can be configured to download different root images.
This is useful both for having various root images for subsets of the \ac{Raspi}s,
but also as a basic way inform the \ac{Raspi}s that the root image changed such
that they can download the new root image at next boot.

As of this writing, there is unfortunately a bug in wget that makes wget not showing
its progress bar while downloading the root image at startup. Instead, it may appear
to be unable to connect although it is downloading, so keep patient.

\begin{lstlisting}[]
Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
image.csv       100% |*******************************************************| 1287   0:00:00 ETA
Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
_
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

REMEMBER TO COPY-PASTE ALL CONTENT FROM SCRIPT INTO LATEX AGAIN IN CASE SOME
SCRIPT CHANGED, BUT WAS NOT UPDATED IN LATEX
